---
title: "Section 6 - ggplot2"
author: "Jen Birstler (with minor revisions by John Gillett)"
output: html_document
---

##Packages

ggplot2 is a "package" for R. The functions for it are not already included in R and have to be installed. For the first time using ggplot2, install it using `install.packages("ggplot2")` then load it in using `require("ggplot2")`. After that, each time you start an R session and want to use ggplot2, call it into the session with `require("ggplot2")`. Those functions, `install.packages()` and `require()`, can be used with any package in R.

```{r}
if (!require("ggplot2")) {
  install.packages("ggplot2")
  stopifnot(require("ggplot2"))
  }

trees <- read.csv("ggplot2_trees.csv", row.names=1)
```

##qplot

`qplot()` means "quick plot." Specify the dataset, the variables for each axis, and the "geometry" of a boxplot.

```{r}
qplot(data = trees, x = species, y = dbh, geom="boxplot")
```

##ggplot

`ggplot()` initializes a ggplot object and gives us a bit more control than qplot (more details below). After the object is initialized, the geometry "layer" can be added to the plot. 

```{r}
ggplot(data = trees, aes(x = species, y = dbh)) + geom_boxplot()
```

##`aes()` and `geom_SOMETHING()`
`aes()` describes the "aesthetics" of the visuzalization. It determines how the data is encoded on the graph. `geom_SOMETHING()` layers the geometry on the graph. Certain geometries require or ignore certain aesthetics.

```{r}
ggplot(data=trees, aes(x=dbh)) + geom_density()
ggplot(data = anscombe, aes(x = x1, y=y1)) + geom_point() + geom_smooth() + geom_line()
ggplot(data = anscombe, aes(x = x1, y=y1)) + geom_point() + geom_smooth(method="lm") + geom_line()
ggplot(data = mtcars, aes(x=mpg, y=qsec, color=factor(cyl), shape=factor(gear))) + geom_point()
```

##`facet_wrap()` and `facet_grid()`
`facet_wrap()` and `facet_grid()` create multiple graphs on the same plot. A formula is passed to create the facets. (In R's formula notation, `.` indicates all other columns in the dataframe)

```{r}
ggplot(data = mtcars, aes(x=mpg, y=qsec, color=factor(cyl))) + geom_point() + facet_grid(.~gear)
ggplot(data = mtcars, aes(x=mpg, y=qsec, color=factor(cyl))) + geom_point() + facet_grid(carb~gear)
ggplot(data = mtcars, aes(x=mpg, y=qsec, color=factor(cyl))) + geom_point() + facet_wrap(carb~gear)
```

##Axes labels, scales, and limits
Axes titles can be added with `xlab()` and `ylab()`. Main titles can be added with `labs(title= "Title Text")`. Scale transformations can be done to (most) variables in `aes()` and the limits can be set. Limits can also be set for x without a transformation by using `scale_x_continuous()` or `scale_x_discrete()`, similarly for y and other aesthetics. 

```{r}
ggplot(data = trees, aes(x = species, y = dbh)) + geom_boxplot() + ylab("Diameter") + xlab("Species") + labs(title="Some trees are bigger than others")
ggplot(data=trees, aes(x=dbh)) + geom_density() + scale_x_sqrt(limits=c(0, 50))
```
